---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE
)
```

# rTAGS

The goal of rTAGS is to make it easy to work with Twitter data collected via [TAGS](https://tags.hawksey.info/) in R. This package makes ample use of the [{rtweet} package](https://rtweet.info/) to process and prepare the data.

## Installation

You can install the released version of rTAGS from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("devtools")
devtools::install_github('bretsw/rTAGS)
```

## Example

This is a basic example which shows you how to read a TAGS sheet and then use {rtweet} (via `pull_data()`) to access additional data.

```{r example, eval = TRUE}
library(rTAGS)
library(tidyverse)

d <- "https://docs.google.com/spreadsheets/d/1WM2xWG9B0Wqn3YG5uakfy_NSAEzIFP2nEAJ5U_fqufc/edit#gid=8743918" %>%
        read_tags() %>% 
        pull_data(n = 1000)
```

Here is the result:

```{r}
glimpse(d)
```

## Creating an edgelist from the TAGS data

```{r}
edgelist <- create_edgelist(d)

edgelist

edgelist %>% 
  count(edge_type)
```

If you want to simply view the TAGS archive, you can use `read_tags()`:

```{r, eval = FALSE}
d1 <- "https://docs.google.com/spreadsheets/d/1WM2xWG9B0Wqn3YG5uakfy_NSAEzIFP2nEAJ5U_fqufc/edit#gid=8743918" %>% 
        read_tags()
```

Here is the result:

```{r, eval = FALSE}
glimpse(d1)
```
